---
title: "evoxploit"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE
                      fig.path = "man/figures/README-")
```

<!-- html_document -->

## Overview

evoxploit is a package for augmenting a **longitudinal dataset**<sup>&ast;</sup> 
with so called **evolution features** (Niemann et al., 2015) and 
**sequence features** (Hielscher et al., 2014). Theses features reflect 
a study participant's individual change over time, his/her change over time 
in comparison with the cluster he/she was closest to at the baseline 
examinations, and the change of whole participant groups. 

<sup>&ast;</sup> The functions of this package require a n*m-dimensional 
dataframe for the input variables and a factor vector of length n 
with the target variable 
(e.g. a medical outcome). The input data frame should have the following 
variables names semantics:

- Each variable name (e.g. `waist_circum_s0`) consists of a _stem_ and a _suffix_. 
The stem is the substring until the last underscore `_` (`wairst_circum`). 
The suffix is the substring beginning after the last underscore (`s0`).
- The suffix indicates the study wave where the measurement was conducted. For 
instance `s0` could represent the baseline examinations of a longitudinal 
study, `s1` the first follow-
up examinations, `s2` the second follow up examinations and so on.
- All variables must have a suffix. The suffix must start with "s" and end with 
a digit.

## Installation

```{r, eval=FALSE}
# Development version from Github
devtools::install_github("unmnn/evoxploit")
```

## Usage Example

<!-- The example data `epi` is a list with two elements. `data` is a dataframe  -->
<!-- containing 19 input features -->

```{r, message=FALSE, warning = FALSE}
# Inspect example data
library(evoxploit)
str(epi)

# Apply clustering
clu <- clustering(epi$data, label = epi$label$label)
str(clu)

# Generate evolution features
df_evo <- create_CBMS15_attributes(epi$data, 
                                   label = epi$label$label,
                                   li_clustering = clu)
str(df_evo)


# Generate sequence features
df_seq <- create_IDA14_attributes(epi$data, label = epi$label$label)
str(df_seq)

# Generate simple features describing a study participant's change over time
df_cha <- create_simple_attributes(epi$data)
str(df_cha)
```


```{r, fig.height=6.5}
# Calculate Gain Ratio of both original and extracted features w.r.t. the 
# target variable
df_combined <- bind_cols(epi$data, df_evo, df_seq, df_cha, epi$label)
gain_names <- FSelector::gain.ratio(label ~ ., data = df_combined)

gain_names %>%
  tibble::rownames_to_column("variable") %>%
  arrange(desc(attr_importance)) %>%
  mutate(variable = forcats::fct_reorder(variable, attr_importance)) %>%
  slice(1:20) %>%
  ggplot(aes(variable, attr_importance)) +
  geom_col() +
  coord_flip()

```



## Bibliography

- (Niemann et al., 2015) Uli Niemann, Tommy Hielscher, Myra Spiliopoulou, 
Henry Völzke, and Jens-Peter Kühn. "Can we classify the participants of a 
longitudinal epidemiological study from their previous evolution?" 
_Proc. of IEEE Computer-Based Medical Systems_, 121-126, 2015. 
- (Hielscher et al. 2014) Tommy Hielscher, Myra Spiliopoulou, Henry Völzke, 
and Jens-Peter Kühn. "Mining longitudinal epidemiological data to understand a 
reversible disorder". _Proc. of Intelligent Data Analysis_, 2014. 
